---
title: "Week 9: Code along 8"
subtitle: "CHAPTER 6 Kurtosis"
author: "Daniel Lee"
date: "2022-09-19"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
# Load packages

# Core
library(tidyverse)
library(tidyquant)

```

## Goal
Measure portfolio risk using kurtosis. It describes the fatness of the tails in probability distributions. In other words, it measures whether a distribution has more or less returns in its tails than the normal distribution. It matters to investors because a distribution with excess kurtosis (kurtosis > 3) means our portfolio might be at risk of a rare but huge downside event. Kurtosis less than 3 means the portfolio is less risky because it has fewer returns in the tails.

five stocks: "SPY", "EFA", "IJS", "EEM", "AGG"

from 2012-12-31 to 2017-12-31

## 1 Import stock prices

```{r}
symbols <- c("SPY", "EFA", "IJS", "EEM", "AGG")

prices <- tq_get(x    = symbols,
                 get  = "stock.prices",    
                 from = "2012-12-31",
                 to   = "2017-12-31")
```

## 2 Convert prices to returns

```{r}
asset_returns_tbl <- prices %>%
    
    group_by(symbol) %>%
    
    tq_transmute(select     = adjusted, 
                 mutate_fun = periodReturn, 
                 period     = "monthly",
                 type       = "log") %>%
    
    slice(-1) %>%
    
    ungroup() %>%
    
    set_names(c("asset", "date", "returns"))
```

## 3 Assign a weight to each asset

```{r}
# symbols
symbols <- asset_returns_tbl %>% distinct(asset) %>% pull()
symbols

# weights
weights <- c(0.25, 0.25, 0.2, 0.2, 0.1)
weights

w_tbl <- tibble(symbols, weights)
w_tbl
```

## 4 Build a portfolio

```{r}
# ?tq_portfolio

portfolio_returns_tbl <- asset_returns_tbl %>%
    
    tq_portfolio(assets_col = asset, 
                 returns_col = returns, 
                 weights = w_tbl, 
                 rebalance_on = "months", 
                 col_rename = "returns")

portfolio_returns_tbl
```

## 5 Calculate Skewness

```{r}
portfolio_returns_tbl %>%

    tq_performance(Ra = returns,
                   Rb = NULL,
                   performance_fun = table.Stats) %>%
    select(Kurtosis)
```


## 6 Plot

### Distribution of portfolio returns
```{r}
portfolio_returns_tbl %>% 
    ggplot(aes(returns)) + 
    geom_histogram()

```


### Expected Return vs Risk
```{r}
# Figure 6.3 Asset and Portfolio Kurtosis Comparison ----

asset_returns_kurtosis_tbl <- asset_returns_tbl %>%

    # kurtosis for each asset
    group_by(asset) %>%
    summarise(kt = kurtosis(returns),
              mean = mean(returns)) %>%
    ungroup() %>%

    # kurtosis of portfolio
    add_row(tibble(asset = "Portfolio",
                   kt = kurtosis(portfolio_returns_tbl$returns),
                   mean = mean(portfolio_returns_tbl$returns)))

asset_returns_kurtosis_tbl %>%

    ggplot(aes(kt, mean)) +
    geom_point() +
    
    # Formatting
    scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
    theme(legend.position = "none") +

    # Add label
    ggrepel::geom_text_repel(aes(label = asset, color = asset), size = 5) +

    labs(y = "Expected Return",
         x = "Kurtosis")

```

### Rolling kurtosis
```{r}

# 3 Rolling kurtosis ----

# Assign a value to winder
window <- 24

port_rolling_kurtosis_tbl <- portfolio_returns_tbl %>%

    tq_mutate(select = returns,
              mutate_fun = rollapply,
              width      = window,
              FUN        = kurtosis,
              col_rename = "rolling_kurtosis") %>%
    select(date, rolling_kurtosis) %>%
    na.omit()

# Figure 6.5 Rolling kurtosis ggplot ----

port_rolling_kurtosis_tbl %>%

    ggplot(aes(date, rolling_kurtosis)) +
    geom_line(color = "cornflowerblue") +

    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    scale_x_date(breaks = scales::breaks_pretty(n = 7)) +

    labs(title = paste0("Rolling ", window, "-Month Kurtosis"),
         x = NULL,
         y = "kurtosis") +
    theme(plot.title = element_text(hjust = 0.5)) +

    annotate(geom = "text",
             x = as.Date("2016-12-01"), y = 3,
             color = "red", size = 5,
             label = str_glue("The risk level skyrocketed at the end of the period
                              with the 24-month kurtosis rising above three."))
```

