---
title: "Week 2: Code along 1"
author: "Daniel Lee"
date: "2022-08-30"
output:
  html_document: 
    toc: yes
  pdf_document: default
  word_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Welcome

# Ch1 Introduction

## The data science project workflow

## Prerequisites

-   R
-   RStudio
-   r packages

## Install the tidyverse package

```{r eval = FALSE}
install.packages("tidyverse")
library(tidyverse)
```

## Running R code

```{r}
1+2
```

## Getting help

* Google 
* Stackoverflow

# Ch2 Introduction to Data Exploration

# Ch3 Data Visualization

## Set up

```{r}
library(tidyverse)
```

## data

```{r}
mpg
```

## aesthetics

* x
* y
* color
* size
* alpha
* shape

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# Add a third variable: color
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))

# Add a third variable: size
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))

# Add a third variable: alpha
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))

# Add a third variable: shape
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))

# Setting aesthetics manually
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

## common problems

* Sometimes you’ll run the code and nothing happens.
* Putting the + in the wrong place.

How to get help

* ? function name
* Select the function name and press F1
* Read the error message
* Google the error message

## facets

```{r}
# facet your plot by a single variable
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)

# facet your plot on the combination of two variables
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```

## geometric objects

different visual object to represent data

```{r}
# scatterplot
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# smooth line
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

not every aesthetic works with every geom

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

two geoms in the same graph!

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv)) +
    
  # Add another geom
  geom_point(mapping = aes(x = displ, y = hwy, color = drv))
```

local vs. global mappings
This makes it possible to display different aesthetics in different layers.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()

```

specify different data for each layer

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```

## statistical transformation

## position adjustments

adjustments for bar charts

```{r}
# to place each object exactly where it falls
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(position = "identity")

# to compare proportions across groups
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(position = "fill")

# to compare individual values
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(position = "dodge")
```

adjustments for scatterplots

```{r}
# overlapping
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point()

# jitter
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_jitter()
```


## coordinate systems

switch x and y

```{r}
# original
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()

# switch x and y
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```

set the aspect ratio correctly for maps

```{r}
nz <- map_data("nz")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()
```

Polar coordinates reveal an interesting connection between a bar chart and a Coxcomb chart.

```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```

## the layered grammar of graphics

The grammar of graphics is based on the insight that you can uniquely describe any plot as a combination of:

* a dataset, 
* a geom, 
* a set of mappings, 
* a stat, 
* a position adjustment, 
* a coordinate system, and 
* a faceting scheme.
