---
title: "Apply it to your data 13"
subtitle: "How much should you expect from your $100 investment after 20 years?"
author: "Daniel Lee"
date: "2022-09-19"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
# Load packages

# Core
library(tidyverse)
library(tidyquant)

# Source function
source("../00_scripts/simulate_accumulation.R")
```


## 1 Import stock prices

***Revise the code below.***

* Replace symbols with your stocks.
* Replace the from and the to arguments to date from 2012-12-31 to present.

```{r}
symbols <- c("SPY", "EFA", "IJS", "EEM", "AGG")

prices <- tq_get(x    = symbols,
                 get  = "stock.prices",    
                 from = "2012-12-31",
                 to   = "2017-12-31")
```

## 2 Convert prices to returns

```{r}
asset_returns_tbl <- prices %>%
    
    group_by(symbol) %>%
    
    tq_transmute(select     = adjusted, 
                 mutate_fun = periodReturn, 
                 period     = "monthly",
                 type       = "log") %>%
    
    slice(-1) %>%
    
    ungroup() %>%
    
    set_names(c("asset", "date", "returns"))
```

## 3 Assign a weight to each asset

***Revise the code for weights.*** 

* The vector weights should have a length equal to the number of assets in the portfolio.
* The values in the vector weights should sum to 1.

```{r}
# symbols
symbols <- asset_returns_tbl %>% distinct(asset) %>% pull()
symbols

# weights
weights <- c(0.25, 0.25, 0.2, 0.2, 0.1)
weights

w_tbl <- tibble(symbols, weights)
w_tbl
```

## 4 Build a portfolio


```{r}
portfolio_returns_tbl <- asset_returns_tbl %>%
    
    tq_portfolio(assets_col = asset, 
                 returns_col = returns, 
                 weights = w_tbl, 
                 rebalance_on = "months", 
                 col_rename = "returns")

portfolio_returns_tbl
```

## 5 Simulating growth of a dollar

```{r}
# Get mean portfolio return
mean_port_return <- mean(portfolio_returns_tbl$returns)
mean_port_return

# Get standard deviation of portfolio returns
stddev_port_return <- sd(portfolio_returns_tbl$returns)
stddev_port_return
```



## 6 Simulation function

***No need***


## 7 Running multiple simulations 

## 8 Visualizing simulations with ggplot
Line Plot of Simulations with Max, Median, and Min

Based on the Monte Carlo simulation results, how much should you expect from your $100 investment after 20 years? What is the best-case scenario? What is the worst-case scenario? What are limitations of this simulation analysis? 

