---
title: "R for Data Science"
author: "Daniel Lee"
date: "2022-09-09"
output:
  html_document: 
    toc: yes
  pdf_document: default
  word_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(nycflights13)
```

# Import data

```{r}
flights
```


# Filter rows

```{r}
filter(flights, month == 1, day == 1)
filter(flights, month == 1 & day == 1)
filter(flights, month == 1 | day == 1)
filter(flights, month %in% c(11, 12))
```

# Arrange rows

```{r}
arrange(flights, desc(month), desc(day))
```

# Select columns

```{r}
select(flights, year:dep_time)
select(flights, year, month, day, dep_time)
select(flights, year, month, day, dep_time, dep_delay)
select(flights, year, month, day, starts_with("dep"))
select(flights, year, month, day, contains("time"))
select(flights, year, month, day, ends_with("time"))
select(flights, year, month, day, contains("time"), everything())

```

# Add columns

```{r}
mutate(flights,
       gain = dep_delay - arr_delay) %>%
    
    # Select year, month, day, and gain
    select(year:day, gain)

# Just keep gain
mutate(flights,
       gain = dep_delay - arr_delay) %>%
    
    # Select year, month, day, and gain
    select(gain)

# alternative using transmute()
transmute(flights,
          gain = dep_delay - arr_delay)

# lag()
select(flights, dep_time) %>%
    
    mutate(dep_time_lag1 = lag(dep_time))

# cumsum()
select(flights, minute) %>%
    
    mutate(minute_cumsum = cumsum(minute))
```

# Summarize by groups

Collapsing data to a single row
```{r}
flights

# average departure delay
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```


Summarize by group
```{r}
flights %>%
    
    # Group by airlines
    group_by(carrier) %>%
    
    # Calculate average departure delay
    summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
    
    # Sort it
    arrange(delay)
```


Delays increase with distance up to ~750 miles and then decrease
```{r}
flights %>%
    group_by(dest) %>%
    summarise(count = n(),
              dist = mean(distance, na.rm = TRUE),
              delay = mean(arr_delay, na.rm = TRUE)) %>%
    
    # Plot
    ggplot(mapping = aes(x = dist, y = delay)) +
    geom_point(aes(size = count), alpha = 0.3) +
    geom_smooth(se = FALSE)
```


Missing values
```{r}
flights %>%
    
    # Remove missing values
    filter(!is.na(dep_delay))
```


counts
Wow, there are some planes that have an average delay of 5 hours (300 minutes)!



We can get more insight if we draw a scatterplot of number of flights vs. average delay:


useful summary functions


grouping multiple variables
```{r}
flights %>%
    group_by(year, month, day) %>%
    summarise(count = n()) %>%
    ungroup()
```


ungrouping













