---
title: "Week 14: Code along 12"
subtitle: "CHAPTER 10 Component Contribution to Standard Deviation"
author: "Daniel Lee"
date: "2022-09-19"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
# Load packages

# Core
library(tidyverse)
library(tidyquant)
library(readr)

# Time series
library(lubridate)
library(tibbletime)

# modeling
library(broom)
```

## Goal
Examine how each asset contributes to portfolio standard deviation. This is to ensure that our risk is not concentrated in any one asset.


five stocks: "SPY", "EFA", "IJS", "EEM", "AGG"
from 2012-12-31 to 2017-12-31

## 1 Import stock prices

```{r}
symbols <- c("SPY", "EFA", "IJS", "EEM", "AGG")

prices <- tq_get(x    = symbols,
                 get  = "stock.prices",    
                 from = "2012-12-31",
                 to   = "2017-12-31")
```

## 2 Convert prices to returns

```{r}
asset_returns_tbl <- prices %>%
    
    group_by(symbol) %>%
    
    tq_transmute(select     = adjusted, 
                 mutate_fun = periodReturn, 
                 period     = "monthly",
                 type       = "log") %>%
    
    slice(-1) %>%
    
    ungroup() %>%
    
    set_names(c("asset", "date", "returns"))
```

## 3 Component Contribution Step-by-Step

Refresh your memory on covariance with this video. Click [this link](https://www.youtube.com/watch?v=qtaqvPAeEJY)
Refresh your memory on matrix multiplication. Click [this link](https://www.mathsisfun.com/algebra/matrix-multiplying.html) 


```{r}
# Transform data into wide form
asset_returns_wide_tbl <- asset_returns_tbl %>%

    pivot_wider(names_from = asset, values_from = returns) %>%

    column_to_rownames(var = "date")

asset_returns_wide_tbl

# Covariance of asset returns
covariance_matrix <- cov(asset_returns_wide_tbl)

covariance_matrix

# Standard deviation of portfolio
# Summarizes how much each asset's returns vary with those of other assets within the portfolio into a single number
w <- c(0.25, 0.25, 0.2, 0.2, 0.1)

sd_portfolio <- sqrt(t(w) %*% covariance_matrix %*% w)
sd_portfolio


# Component contribution
# Similar to the formula for sd_portfolio
# Mathematical trick to summarize the same, sd_portfolio, by asset instead of a single number
component_contribution <- (t(w) %*% covariance_matrix * w) / sd_portfolio[1,1]
component_contribution

rowSums(component_contribution)

# Component contribution in percentage
component_percentages <- (component_contribution / sd_portfolio[1,1]) %>%
    round(3) %>%
    as_tibble()

component_percentages

component_percentages %>%

    as_tibble() %>%
    gather(key = "asset", value = "contribution")
```


## 4 Component Contribution with a Custom Function



## 5 Visualizing Component Contribution



## 6 Rolling Component Contribution






