---
title: "Week 10: Code along 9"
subtitle: "CHAPTER 7 Sharpe Ratio"
author: "Daniel Lee"
date: "2022-09-19"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
# Load packages

# Core
library(tidyverse)
library(tidyquant)

```

## Goal
Measure which portfolio is expected to yield a higher return adjusted for risk using the Sharpe Ratio.
The Sharpe Ratio is the mean of the excess portfolio returns (returns - risk free rate) divided by the standard deviation of the excess returns.

five stocks: "SPY", "EFA", "IJS", "EEM", "AGG"

from 2012-12-31 to 2017-12-31

## 1 Import stock prices

```{r}
symbols <- c("SPY", "EFA", "IJS", "EEM", "AGG")

prices <- tq_get(x    = symbols,
                 get  = "stock.prices",    
                 from = "2012-12-31",
                 to   = "2017-12-31")
```

## 2 Convert prices to returns

```{r}
asset_returns_tbl <- prices %>%
    
    group_by(symbol) %>%
    
    tq_transmute(select     = adjusted, 
                 mutate_fun = periodReturn, 
                 period     = "monthly",
                 type       = "log") %>%
    
    slice(-1) %>%
    
    ungroup() %>%
    
    set_names(c("asset", "date", "returns"))
```

## 3 Assign a weight to each asset

```{r}
# symbols
symbols <- asset_returns_tbl %>% distinct(asset) %>% pull()
symbols

# weights
weights <- c(0.25, 0.25, 0.2, 0.2, 0.1)
weights

w_tbl <- tibble(symbols, weights)
w_tbl
```

## 4 Build a portfolio

```{r}
# ?tq_portfolio

portfolio_returns_tbl <- asset_returns_tbl %>%
    
    tq_portfolio(assets_col = asset, 
                 returns_col = returns, 
                 weights = w_tbl, 
                 rebalance_on = "months", 
                 col_rename = "returns")

portfolio_returns_tbl
```

## 5 Calculate Sharpe Ratio

```{r}
# Risk free rate
rfr <- 0.0003

portfolio_sharpe_tbl <- portfolio_returns_tbl %>%

    tq_performance(Ra = returns,
                   Rf = rfr,
                   performance_fun = SharpeRatio,
                   FUN = "StdDev") 

portfolio_sharpe_tbl
```


## 6 Plot

### Returns Histogram with Risk-Free Rate
```{r}
# Figure 7.2 Returns Histogram with Risk-Free Rate ggplot ----

portfolio_returns_tbl %>%

    ggplot(aes(returns)) +
    geom_histogram(binwidth = 0.01, fill = "cornflowerblue", alpha = 0.5) +
    geom_vline(xintercept = rfr, color = "green", size = 1) +

    annotate(geom= "text",
             x = rfr + 0.002, y = 13,
             label = "risk free rate", angle = 90, size = 5) +
    labs(y = "count")

```


### Scatter Returns around Risk Free Rate
```{r}
# Figure 7.1 Scatter Returns around Risk Free Rate ----

portfolio_returns_tbl %>%

    # Transform data
    mutate(returns_excess = if_else(returns > rfr, "above_rfr", "below_rfr")) %>%

    ggplot(aes(date, returns, color = returns_excess)) +
    geom_point(show.legend = FALSE) +

    # risk free rate
    geom_hline(yintercept = rfr, linetype = "dotted", size = 1, color = "cornflowerblue") +

    # election date
    geom_vline(xintercept = as.Date("2016-11-30"), size = 1, color = "cornflowerblue") +

    # formatting
    scale_x_date(breaks = scales::pretty_breaks(n = 7)) +

    # labeling
    annotate(geom = "text",
             x = as.Date("2017-01-01"), y = -0.04,
             label = "Election", angle = 90, size = 5) +
    annotate(geom = "text",
             x = as.Date("2017-06-01"), y = -0.01,
             label = str_glue("No returns below the RFR
                              after the 2016 election"),
             color = "red", size = 4) +
    labs(y = "percent monthly returns",
         x = NULL)


```


### Rolling Sharpe
```{r}
# Custom function
# necessary because we would not be able to specify FUN = "StdDev" otherwise

calculate_rolling_sharpeRatio <- function(df) {

    SharpeRatio(df,
                Rf = rfr,
                FUN = "StdDev")

}

# dump(list = "calculate_rolling_sharpeRatio",
#      file = "00_scripts/calculate_rolling_sharpeRatio.R")

# Set the length of periods for rolling calculation
window <- 24

# Calculate rolling sharpe ratios
rolling_sharpe_tbl <- portfolio_returns_tbl %>%

    tq_mutate(select = returns,
              mutate_fun = rollapply,
              width = window,
              align = "right",
              FUN = calculate_rolling_sharpeRatio,
              col_rename = "sharpeRatio") %>%
    na.omit()

rolling_sharpe_tbl


# Figure 7.5 Rolling Sharpe ggplot ----

rolling_sharpe_tbl %>%

    ggplot(aes(date, sharpeRatio)) +
    geom_line(color = "cornflowerblue") +

    labs(title = paste0("Rolling ", window, "-Month Sharpe Ratio"),
         y = "rolling Sharpe Ratio",
         x = NULL) +
    theme(plot.title = element_text(hjust = 0.5)) +

    annotate(geom = "text",
             x = as.Date("2016-06-01"), y = 0.5,
             label = "This portfolio has done quite well since 2016.",
             size = 5, color = "red")

```

