---
title: "Week 12: Apply it to your data 11"
subtitle: "Chapter 19 Functions"
author: "Daniel Lee"
date: "2022-10-05"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load package
library(tidyverse)
library(nycflights13)
```

## Import your data

```{r}
data(flights)

flights %>% skimr::skim()
```


## Create Data frame functions

### Example 1: count columns

#### code snippets
```{r}
ncol_num <- flights %>%
    
    # Select a type of variables
    select(where(is.numeric)) %>%
    
    # Count columns
    ncol()

ncol_num
```

#### Turn them into a function
```{r}
# Create a function to count # of a type of columns
count_numeric_var <- function(.data) {
    
    # body
    ncol_num <- .data %>%
    
    # Select a type of variables
    select(where(is.numeric)) %>%
    
    # Count columns
    ncol()
    
    # return value
    return(ncol_num)
    
}

flights %>% count_numeric_var()
flights %>% .[, -1:-13] %>% count_numeric_var()
```


#### Adding arguments for details of operation
```{r}
# Create a function to count # of a type of columns
count_type_of_var <- function(.data, type = "numeric") {
    
    # if statement for type of variables
    if(type == "numeric") {
        
        # body
        ncol_num <- .data %>%
        
        # Select a type of variables
        select(where(is.numeric)) %>%
        
        # Count columns
        ncol()
        
    } else if(type == "character") {
        
        # body
        ncol_num <- .data %>%
        
        # Select a type of variables
        select(where(is.character)) %>%
        
        # Count columns
        ncol()        
        
    }
    

    # return value
    return(ncol_num)
    
}

flights %>% count_type_of_var(type = "character")
flights %>% .[, -1:-13] %>% count_type_of_var(type = "character")
```

### Example 2: count rows

#### code snippets
```{r}
nrow_num <- flights %>%
    
    # filter rows that meet a condition
    filter(carrier == "UA") %>%
    
    # Count rows
    nrow()

nrow_num
```

#### Turn them into a function
```{r}
# Create a function to count # of a type of columns
count_n_flights_by_carrier <- function(.data, carrier_name) {
    
    # body
    nrow_num <- .data %>%
        
        # filter rows that meet a condition
        filter(carrier == carrier_name) %>%
        
        # Count rows
        nrow()
    
    nrow_num
    
    # return value
    return(nrow_num)
    
}

flights %>% count_n_flights_by_carrier(carrier_name = "UA")
flights %>% .[1:10, ] %>% count_n_flights_by_carrier(carrier_name = "UA")
```

### Example 3: count rows
Create your own.

#### code snippets
Use the filter() function to select rows that meet a condition.
Refer to Chapter 5.2 Filter rows with filter()

#### Turn them into a function
