---
title: "Week 12: Apply it to your data 11"
subtitle: "Chapter 19 Functions"
author: "Daniel Lee"
date: "2022-10-05"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load package
library(tidyverse)
library(nycflights13)
```

## Import your data

```{r}
data(flights)

flights %>% skimr::skim()
```


## Create Data frame functions

### Example 1: count columns

#### code snippets
```{r}
ncol_num <- flights %>%
    
    # Select a type of variables
    select(where(is.numeric)) %>%
    
    # Count columns
    ncol()

ncol_num
```

#### Turn them into a function



#### Adding arguments for details of operation


### Example 2: count rows

#### code snippets
```{r}
nrow_num <- flights %>%
    
    # filter rows that meet a condition
    filter(carrier == "UA") %>%
    
    # Count rows
    nrow()

nrow_num
```

#### Turn them into a function


### Example 3: count rows
Create your own.

#### code snippets
Use the filter() function to select rows that meet a condition.
Refer to Chapter 5.2 Filter rows with filter()

#### Turn them into a function
