---
title: "Week 11: Code along 10"
subtitle: "CHAPTER 8 Capital Asset Pricing Model"
author: "Daniel Lee"
date: "2022-09-19"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
# Load packages

# Core
library(tidyverse)
library(tidyquant)

```

## Goal
Measure the portfolio's beta coefficient, which can be thought of as the portfolio's sensitivity to the market or its riskiness relative to the market.

five stocks: "SPY", "EFA", "IJS", "EEM", "AGG"

market: "SPY"

from 2012-12-31 to 2017-12-31

## 1 Import stock prices

```{r}
symbols <- c("SPY", "EFA", "IJS", "EEM", "AGG")

prices <- tq_get(x    = symbols,
                 get  = "stock.prices",    
                 from = "2012-12-31",
                 to   = "2017-12-31")
```

## 2 Convert prices to returns

```{r}
asset_returns_tbl <- prices %>%
    
    group_by(symbol) %>%
    
    tq_transmute(select     = adjusted, 
                 mutate_fun = periodReturn, 
                 period     = "monthly",
                 type       = "log") %>%
    
    slice(-1) %>%
    
    ungroup() %>%
    
    set_names(c("asset", "date", "returns"))
```

## 3 Assign a weight to each asset

```{r}
# symbols
symbols <- asset_returns_tbl %>% distinct(asset) %>% pull()
symbols

# weights
weights <- c(0.25, 0.25, 0.2, 0.2, 0.1)
weights

w_tbl <- tibble(symbols, weights)
w_tbl
```

## 4 Build a portfolio


```{r}
# ?tq_portfolio

portfolio_returns_tbl <- asset_returns_tbl %>%
    
    tq_portfolio(assets_col = asset, 
                 returns_col = returns, 
                 weights = w_tbl, 
                 rebalance_on = "months", 
                 col_rename = "returns")

portfolio_returns_tbl
```



## 5 Calculate CAPM Beta

### 5.1 Get market returns
```{r}
# Get market returns
market_returns_tbl <- tq_get("SPY",
                             get = "stock.prices",
                             from = "2012-12-31",
                             to = "2017-12-31") %>%

    # Convert prices to returns
    tq_transmute(select     = adjusted, 
                 mutate_fun = periodReturn, 
                 period     = "monthly",
                 type       = "log", 
                 col_rename = "returns") %>%
    
    slice(-1)
```

### 5.2 Join returns 
```{r}
# Combine market returns with portfolio returns
portfolio_market_returns_tbl <- portfolio_returns_tbl %>%

    # Add market returns
    mutate(market_returns = market_returns_tbl %>% pull(returns))

```

### 5.3 CAPM Beta
```{r}
# 3 Calculating CAPM Beta ----

# A complete list of functions for performance_fun()
# tq_performance_fun_options()

portfolio_market_returns_tbl %>%

    tq_performance(Ra = returns,
                   Rb = market_returns,
                   performance_fun = CAPM.beta)
```


## 6 Plot

### Scatter with regression line
```{r}
# Figure 8.2 Scatter with regression line from ggplot ----

portfolio_market_returns_tbl %>%

    ggplot(aes(market_returns, returns)) +
    geom_point(color = "cornflowerblue") +

    geom_smooth(method = "lm", se = FALSE,
                size = 1.5, color = tidyquant::palette_light()[3]) +

    labs(x = "market returns",
         y = "portfolio returns")

```

### Actual versus fitted returns
```{r}
# Figure 8.5 Actual versus fitted returns ----

portfolio_market_returns_tbl %>%

    # Run regression
    lm(returns ~ market_returns, data = .) %>%

    # Get fitted
    broom::augment() %>%

    # Add date %>%
    mutate(date = portfolio_market_returns_tbl$date) %>%

    # Transform data to long format
    pivot_longer(cols = c(returns, .fitted),
                 names_to = "type",
                 values_to = "returns") %>%

    # Plot
    ggplot(aes(date, returns, color = type)) +
    geom_line()


```

