---
title: "Apply it to your data 6"
subtitle: "How does your portfolio expect perform regarding expected returns and risk?"
author: "Daniel Lee"
date: "2022-09-19"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
# Load packages

# Core
library(tidyverse)
library(tidyquant)

```

## Goal
Visualize expected returns and risk to make it easier to compare the performance of multiple assets and portfolios.

Choose your stocks.

from 2012-12-31 to 2017-12-31

## 1 Import stock prices


## 2 Convert prices to returns (monthly)


## 3 Assign a weight to each asset (change the weigting scheme)


## 4 Build a portfolio


## 5 Plot: Expected Returns versus Risk

How should you expect your portfolio to perform relative to its assets in the portfolio? Would you invest all your money in any of the individual stocks instead of the portfolio? Discuss both in terms of expected return and risk.  
